---
title: "Chapter 3"
author: "Nick Lauerman"
date: "13 February 2020"
output: 
  pdf_document: 
    fig_caption: yes
    fig_height: 6.5
    fig_width: 6.5
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(cache.path = "./Cache/")
```

# Libraries and data used
## Libraries
```{r, cache=FALSE}
library(UserNetR)
library(statnet)
#library(igraph) will be loaded in flow to prevent interferance with statnet
library(intergraph)
```

## Data
```{r}
data("ICTS_G10")
data("DHHS")
```

# Creating and managing networks
# creating net in netstat
```{r}
netmat1 <- rbind(c(0,1,1,0,0),
                 c(0,0,1,1,0),
                 c(0,1,0,0,0),
                 c(0,0,0,0,0),
                 c(0,0,1,0,0))
rownames(netmat1) <- c("A","B","C","D","E")
colnames(netmat1) <- c("A","B","C","D","E")
net1 <- network(netmat1, matrix.type = "adjacency")
class(net1)
summary(net1)
```
```{r, cache=FALSE}
gplot(net1,
      vertex.col = 2,
      displaylabels = TRUE)
```
```{r}
netmat2 <- rbind(c(1,2),
                 c(1,3),
                 c(2,3),
                 c(2,4),
                 c(3,2),
                 c(5,3))
net2 <- network(netmat2,
                matrix.type = "edgelist")
network.vertex.names(net2) <- c("A","B","C","D","E")
summary(net2)
as.sociomatrix(net1)
class(as.sociomatrix(net1))
all(as.matrix(net1) == as.sociomatrix(net1))
as.matrix(net1, matrix.type = "edgelist")
```

## Managing Node and Tie Attribures
### Node attrib
```{r}
set.vertex.attribute(net1, 
                     "gender",
                     c("F","F","M","F","M"))
net1 %v% "alldeg" <- degree(net1)
list.vertex.attributes(net1)
summary(net1)
get.vertex.attribute(net1, "gender")
net1 %v% "alldeg"
```

### Tie Attrib
```{r}
list.edge.attributes(net1)
set.edge.attribute(net1,
                   "rndval",
                   runif(network.size(net1),0,1))
list.edge.attributes(net1)
summary(net1 %e% "rndval")
summary(get.edge.attribute(net1, "rndval"))
netval1 <- rbind(c(0,2,3,0,0),
                 c(0,0,3,1,0),
                 c(0,1,0,0,0),
                 c(0,0,0,0,0),
                 c(0,0,2,0,0))
netval1 <- network(netval1,
                   matrix.type = "adjacency",
                   ignore.eval = FALSE,
                   names.eval = "like")
network.vertex.names(netval1) <- c("A","B","C","D","E")
list.edge.attributes(netval1)
get.edge.attribute(netval1, "like")
as.sociomatrix(netval1)
as.sociomatrix(netval1)
```

## Creating network object in igraph
```{r, cache=FALSE}
detach(package:statnet, unload = TRUE)
detach(package:sna, unload = TRUE)
detach(package:tsna, unload = TRUE)
detach(package:ergm.count, unload = TRUE)
detach(package:tergm, unload = TRUE)
detach(package:networkDynamic, unload = TRUE)
detach(package:ergm, unload = TRUE)
detach(package:network, unload = TRUE)
detach(package:statnet.common, unload = TRUE)
library(igraph)
```
```{r}
inet1 <- graph.adjacency(netmat1)
class(inet1)
summary(inet1)
#str(inet1)
inet2 <- graph.edgelist(netmat2)
summary(inet2)
V(inet2)$name <- c("A","B","C","D","E")
E(inet2)$val <- c(1:6)
summary(inet2)
#str(inet2)
```

## Going back and forth between statman and igraph
```{r, cache=FALSE}
library(intergraph)
```
```{r}
class(net1)
net1igraph <- asIgraph(net1)
class(net1igraph)
#str(net1igraph)
```

# importing network data
```{r, cache=FALSE}
detach(package:igraph, unload = TRUE)
library(statnet)
```
```{r}
netmat3 <- rbind(c("A","B"),
                 c("A","C"),
                 c("B","C"),
                 c("B","D"),
                 c("C","B"),
                 c("E","C"))
net.df <- data.frame(netmat3)
net.df
write.csv(net.df,
          file = "MyData.csv",
          row.names = FALSE)
net.edge <- read.csv(file = "MyData.csv")
net_import <- network(net.edge,
                      matrix.type = "edgelist")
summary(net_import)
gden(net_import)
```

# Common network data tasks
## Filtering networks daved on vertex or edge attribute values
### Filtering based on Node Values
```{r}
n1F <- get.inducedSubgraph(net1,
                           which(net1 %v% "gender" == "F"))
n1F[,]
```

```{r test plot, cache=FALSE}
gplot(n1F, displaylabels = TRUE)
```

```{r test}
deg <- net1 %v% "alldeg"
n2 <- get.inducedSubgraph(net1, which(deg > 1))
```

```{r, cache=FALSE}
gplot(n2, displaylabels = TRUE)
```

### remove isolates
```{r, cache=FALSE}
data("ICTS_G10")
summary(ICTS_G10, print.adj = FALSE )
```
```{r}
gden(ICTS_G10)
length(isolates(ICTS_G10))
n3 <- ICTS_G10
delete.vertices(n3, isolates(n3))
gden(n3)
length(isolates(n3))
```

### filtering based on edge values
```{r, cache=FALSE}
data("DHHS")
summary(DHHS, print.adj = FALSE )
```
```{r}
d <- DHHS
gden(d)
op <- par(mar = rep(0,4))
gplot(d,
      gmode = "graph",
      edge.lwd = d %e% 'collab',
      edge.col = "grey50",
      vertex.col = "lightblue",
      vertex.cex = 1.0,
      vertex.sides = 20)
par(op)
as.sociomatrix(d)[1:6,1:6]
table(d %e% "collab")
d.val <- as.sociomatrix(d, attrname = "collab")
d.val[d.val < 3] <-  0
d.filt <- as.network(d.val,
                     directed = FALSE,
                     matrix.type = "a",
                     ignore.eval = FALSE,
                     names.eval = "collab")
summary(d.filt, print.adj = FALSE)
gden(d.filt)
op <- par(mar = rep(0,4))
gplot(d.filt,
      gmode = "graph",
      displaylabels = TRUE,
      vertex.col = "lightblue",
      vertex.cex = 1.3,
      label.cex = 0.4,
      label.pos = 5,
      displayisolates = FALSE)
par(op)
op <- par(mar = rep(0,4))
d.val <- as.sociomatrix(d, attrname = "collab")
gplot(d.val,
      gmode = "graph",
      thresh = 2,
      displaylabels = TRUE,
      vertex.col = "lightblue",
      vertex.cex = 1.3,
      label.cex = 0.4,
      label.pos = 5,
      displayisolates = FALSE)
par(op)
```

## transformining directed to non-directed network
```{r}
net1mat <- symmetrize(net1,
                      rule = "weak")
net1mat
net1symm <- network(net1mat,
                    matrix.type = "adjacency")
network.vertex.names(net1symm) <- c("A","B","C","D","E")
summary(net1symm)
```


